---
title: "Exercise 17"
author: "Cameron Bale"
format: docx
---

1. Simulate data like we did in class but with 5 clusters and three variables: average number of days between purchases (`purchase_frequency`), total number of past purchases (`purchase_total`), and membership in a premium subscription program (`premium`).
2. Tune `k` where `k = 1:10` and you cluster just on `purchase_frequency` and `purchase_total`.
3. Produce a scree plot to compare model fit and choose `k`.
4. Use `augment()` to append the predicted cluster using the best-fitting model and produce a cluster profile as a grouped summary for `purchase_frequency`, `purchase_total`, and `premium`.
5. Try and name each cluster based on the cluster profile.
6. Render the Quarto document into Word and upload to Canvas.

**Five points total, one point each for:**

- **Simulating data with 5 clusters and three variables, one being discrete.**
- **Tuning k by running 10 models clustering on just two variables.**
- **Creating a scree plot to compare model fit and select k.**
- **Producing a cluster profile and naming each cluster.**
- **Submitting a rendered Word document.**

NOTE: The solution shown in this file follow the above instructions. However, if you chose to complete a cluster analysis on a data set of your choosing, use the following point breakdown:

**Five points total, one point each for:**

- **Import and clean the data.**
- **Tune k by running multiple clustering models on the variables you chose.**
- **Create a scree plot to compare model fit and select k.**
- **Producing a cluster profile and name each cluster.**
- **Submitting a rendered Word document.**

## Simulate Data

```{r}
# Load packages.
library(tidyverse)
library(tidymodels)

# Simulate centroids.
centroids <- tibble(
  cluster = factor(1:5), 
  purchase_frequency_mean = c(5, 0, 12, 2, 7),
  purchase_total_mean = c(1, 3, 7, 10, 12),
  premium_prob = c(0.75, 0.10, 0.50, 0.80, 0.02),
  num_obs = c(100, 150, 50, 200, 75)
)

# Simulate observations.
set.seed(42)
sim_data <- centroids |>
  mutate(
    purchase_frequency = pmap(list(n = num_obs, mean = purchase_frequency_mean), rnorm),
    purchase_total = pmap(list(n = num_obs, mean = purchase_total_mean), rnorm),
    premium = pmap(list(n = num_obs, size = 1, p = premium_prob), rbinom)
  ) |> 
  unnest(c(purchase_frequency, purchase_total, premium))

sim_data
```

## Hyperparameter Tuning

```{r}
# Select the clustering variables.
sim_obs <- sim_data |> 
  select(purchase_frequency, purchase_total)

# Tune k.
set.seed(42)
fit_tune <- tibble(k = 1:10) |>
  mutate(
    fit_km = pmap(list(centers = k), kmeans, x = sim_obs, nstart = 10),
    model_fit = map(fit_km, glance)
  ) |> 
  unnest(model_fit)

# Scree plot.
ggplot(fit_tune, aes(x = k, y = tot.withinss)) +
  geom_point() +
  geom_line()
```

We know from simulating the data that `k = 5`, and the scree plot agrees.

## Cluster Profile

```{r}
# Fit a k-means model with k = 5.
set.seed(42)
fit_km5 <- sim_data |> 
  select(purchase_frequency, purchase_total) |> 
  kmeans(centers = 5, nstart = 10)

# Append predicted clusters.
sim_data_km5 <- augment(fit_km5, sim_data)

# Produce a cluster profile.
sim_data_km5 |> 
  group_by(.cluster) |> 
  summarize(
    size = n(),
    frequency_mean = mean(purchase_frequency),
    total_mean = mean(purchase_total),
    premium_prop = mean(premium)
  )
```

We use the cluster profile to name each profile. Here's one possible interpretation.

1. **High Rollers**: These customers are often premium members that have made a large number of frequent purchases.
2. **Slow Users**: These customers have made a moderate number of purchases and are more likely than not to be premium subscribers, although they purchase significantly less frequently than any other segment.
3. **Newbies**: Customers who haven't made many purchases but are often premium members.
4. **Dipping Their Toes In**: Customers who are relatively unlikely to be premium members who make frequent purchases but haven't made many in total.
5. **Value Kings/Queens**: Customers who almost never subscribe to premium, but make very large, relatively frequent purchases.

