---
title: "Exercise 21"
author: "Marc Dotson"
format: docx
---

1. Simulate data like we did in class but with 5 clusters and an additional discrete variable `z`.
2. Tune `k` where `k = 1:10` and you cluster just on `x` and `y`.
3. Produce a scree plot to compare model fit and choose `k`.
4. Use `augment()` to append the predicted cluster using the best-fitting model and produce a cluster profile as a grouped summary for `x`, `y`, and `z`.
5. Try and name each cluster based on the cluster profile.
6. Render the Quarto document into Word and upload to Canvas.

**Five points total, one point each for:**

- **Simulating data with 5 clusters and x, y, and z where z is discrete.**
- **Tuning k by running 10 models clustering on just x and y.**
- **Creating a scree plot to compare model fit and select k.**
- **Producing a cluster profile for x, y, and z and naming each cluster.**
- **Submitting a rendered Word document.**

## Simulate Data

```{r}
# Load packages.
library(tidyverse)
library(tidymodels)

# Simulate centroids.
centroids <- tibble(
  cluster = factor(1:5), 
  x_mean = c(5, 0, -3, 2, 7),
  y_mean = c(-1, 1, -2, 5, 9),
  z_prob = c(0.75, 0.10, 0.50, 0.80, 0.02),
  num_obs = c(100, 150, 50, 200, 75)
)

# Simulate observations.
set.seed(42)
sim_data <- centroids |>
  mutate(
    x = pmap(list(n = num_obs, mean = x_mean), rnorm),
    y = pmap(list(n = num_obs, mean = y_mean), rnorm),
    z = pmap(list(n = num_obs, size = 1, p = z_prob), rbinom)
  ) |> 
  unnest(c(x, y, z))

sim_data
```

## Hyperparameter Tuning

```{r}
# Select the clustering variables.
sim_obs <- sim_data |> 
  select(x, y)

# Tune k.
set.seed(42)
fit_tune <- tibble(k = 1:10) |>
  mutate(
    fit_km = pmap(list(centers = k), kmeans, x = sim_obs),
    model_fit = map(fit_km, glance)
  ) |> 
  unnest(model_fit)

# Scree plot.
ggplot(fit_tune, aes(x = k, y = tot.withinss)) +
  geom_point() +
  geom_line()
```

We know from simulating the data that `k = 5` and the scree plot agrees, with a clear elbow at `k = 5`.

## Cluster Profile

```{r}
# Fit a k-means model with k = 5.
set.seed(42)
fit_km5 <- sim_data |> 
  select(x, y) |> 
  kmeans(centers = 5)

# Append predicted clusters.
sim_data_km5 <- augment(fit_km5, sim_data)

# Produce a cluster profile.
sim_data_km5 |> 
  group_by(.cluster) |> 
  summarize(
    size = n(),
    x_mean = mean(x),
    y_mean = mean(y),
    z_prop = mean(z)
  )
```

We use the cluster profile to name each profile. Here's my take, without assigning any meaning to `x`, `y`, and `z`.

1. **Riding High**: This cluster is high on `x` and `y` but lowest on `z`.
2. **Y and Z**: This cluster is second highest on `y` and the highest on `z`.
3. **Nothing**: This cluster is closest to zero across all three variables.
4. **X Please**: This cluster is distinguished by being highest on `x` only.
5. **Negative**: This cluster has negative values for both `x` and `y` and average `z`.
